---
output: html_document
---

# ***Aspects of Round Goby Population Ecology in the Upper St. Lawrence River***
## **Anna L. Haws, John M. Farrell**
## **Mortality Analysis**

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(magrittr)
library(plyr)
library(tidyr)
library(FSA)
library(dplyr)
library(ggplot2)
library(gt)
library(ggpubr)
```

```{r}
# Load fully processed dataset
RG <- read_excel("~/GitHub/Anna_Haws_Traits_of_Round_Goby_in_USLR/Data/DISSECTED.xlsx") 

#Remove unaged and immature data rows
RG %<>% filter(!is.na(RG$Age))
RG %<>% filter(Sex != "I")

#Create Age only count table
RG %<>% group_by(Age)
RGCounts_A<- RG %>% dplyr::count(Age, sort = FALSE)

#Create AGe*Sex count table
RG %<>% group_by(Age, Sex)
RGCounts<- RG %>% dplyr::count(Age, Sex, sort = TRUE)


```

#### A=Annual mortality

#### S= Annual Survival

#### Z=instantaneous mortality rate

$$\ A+S=1 $$ $$\ A=\frac{C_{t}-C_{t+1}}{C_{t}}=1-\frac{C_{t+1}}{C_{t}}$$ $$\ Z=log(C_{t})-log(C_{t+1})$$ $$\ A=1-e^{-Z} $$

## Linear regression fitting, survival estimate, and statistical comparisons between sexes

```{r, echo=TRUE}
#Create df of counts after full recruitment to collection gear
RGCounts_3<- filter(RGCounts, Age>=3)
RGCounts_3$lnct<- log(RGCounts_3$n, base = exp(1))

#Evaluate difference in slope parameter between sexes
lmCC_sex<- lm(lnct~Age*Sex, data=RGCounts_3)
summary(lmCC_sex)

#Female-only regression
RGCounts_3_F<- filter(RGCounts_3, Sex=="F")
lmCC_F<- lm(lnct~Age, data=RGCounts_3_F)
summary(lmCC_F)

#Male-only regression
RGCounts_3_M<- filter(RGCounts_3, Sex=="M")
lmCC_M<- lm(lnct~Age, data=RGCounts_3_M)
summary(lmCC_M)

#All round gobies combined
RGCounts_A_3<- filter(RGCounts_A, Age>=3)
RGCounts_A_3$lnct<- log(RGCounts_A_3$n, base = exp(1))

lmCC_A<- lm(lnct~Age, data=RGCounts_A_3)
summary(lmCC_A)
```




## Chapman-Robson Metric estimation and catch curve

Estimates Annual Survival Rate using:

$$\widehat{S}= \frac{T}{n+T-1}=\frac{\overline{T}}{1+\overline{T}-\frac{1}{n}}$$

### All gobies combined
```{r, echo=TRUE}
#Obtain parameter  (S, Z) estimates using Chapman-Robson function from "FSA" Package
RG_cr <- chapmanRobson(n~Age,data=RGCounts_A,ages2use=3:5)
Measure<- c("S", "Z")
AllRG_CR<-cbind(summary(RG_cr, digits=2), confint(RG_cr, digits=2))
(AllRG_CR<- data.frame(Measure, AllRG_CR))

#Catuch-curve plot
RG_CC<- catchCurve(n~Age, data=RGCounts_A, ages2use =3:5, weighted = TRUE)
plot(RG_CC, pos.est = "bottomleft")
Measure<- c("Z", "A")
(RG_CC<- cbind(summary(RG_CC, digits=2), confint(RG_CC, digits=2)))
(AllRG_CC<- data.frame(Measure, RG_CC))
AllRG_CC<- AllRG_CC[2,]
AllRG_CC<- AllRG_CC[-c(4,5)]

(AllRG_Mortality_table<- rbind(AllRG_CR, AllRG_CC))
```



#### Plot

```{r, echo=FALSE}

(Mortality_plot<-ggplot(data=RGCounts_3, aes(x=Age, y=log(n), colour=Sex)) + geom_point(data=RGCounts_3, aes(x=Age, y=log(n), colour=Sex)) + ylim(0,5)  + scale_colour_manual(values=c("black", "grey")) + labs(x= "Age", y="Log(Catch [n])")+ theme_classic() + theme( legend.position = "bottom", text=element_text(size=12), legend.margin=margin(-10, 0, 0, 0)) + stat_smooth( method = lm, se=FALSE, aes(colour=Sex), data = RGCounts_3) + stat_regline_equation(inherit.aes = TRUE, 
  show.legend = NULL,  label.x = c(1,1), label.y = c(0.25,0.75), aes(label =  paste(..eq.label.., ..adj.rr.label.., sep = "~~~~"))) + geom_point(data=RGCounts, aes(x=Age, y=log(n), colour=Sex)))

```

```{r, echo=FALSE, include=FALSE}
Mortality_plot %>% ggsave(filename="Mortality_plot.tiff", device = "tiff", path="~/GitHub/Anna_Haws_Traits_of_Round_Goby_in_USLR/Figures/TIFFs", dpi="retina", width = 90, height = 90, units="mm")
```
